---
title: "Homework 5"
author: "Ryan"
date: "April 14, 2018"
output: pdf_document
---

```{r, include=FALSE}
library(alr4)
library(car)
attach(UN11)
```
# 6.1
```{r}
NH <- lm(lifeExpF ~ 1)
AH <- lm(lifeExpF ~ group)
anova(NH, AH)
```
The F-statistic of ~160 and corrosponding p-value of nearly 0 implies that group should be included in the model.

# 6.2
Neither model 6.7 or 6.8 are submodels of the other. As such, it would not make sense to do an F-test between them.

# 6.4
## 1
The NH model does not include the main effect from group, however, it does account for group's interaction with log(ppgdp). Since group is not included but a higher-order term including group is, this is a violation of the marginality principle.

##2
```{r}
NH <- lm(lifeExpF ~ log(ppgdp) + group:log(ppgdp))
AH <- lm(lifeExpF ~ group + log(ppgdp) + group:log(ppgdp))
anova(NH, AH)
```
The F-statistic's p-value of 0.056 for group is not significant. As such, we conclude that group should not be included in the model. 

# 6.7
## 1
```{r, include = FALSE}
attach(fuel2001)
```
```{r}
m1 <- lm(FuelC ~ Tax + Drivers + Income + log(Miles))
m2 <- lm(FuelC ~ log(Miles) + Income + Drivers + Tax)
anova(m1)
anova(m2)
```
According to Type I ANOVA on the first model, the predictors Tax and Drivers are the only significant ones. As for Type I ANOVA on the second model, log(Miles), Income and Drivers are all significant. 

## 2
```{r}
Anova(m1, type = "II")
Anova(m2, type = "II")
```
The Type II ANOVA test are the same as the the Type I ANOVA test on the first model (Since Drivers is far and away the most significant term.)

# 6.8
attached image

# 6.9
## 1
```{r, include = FALSE}
attach(cakes)
```
```{r}
NH <- lm(Y ~ X1 + I(X1^2) + X2 + I(X2^2))
AH <- lm(Y ~ X1 + I(X1^2) + X2 + I(X2^2) + X1:X2)
anova(NH, AH)
```
The F-statistic of 15.1 and corrosponding p-value well under 0.05 implies that the model without the interaction term perfroms significantly worse. As such, we conclude that X1:X2 should not be dropped. 
```{r}
NH <- lm(Y ~ X1 + X2 + I(X2^2) + X1:X2)
AH <- lm(Y ~ X1 + I(X1^2) + X2 + I(X2^2) + X1:X2)
anova(NH, AH)
```
The F-statistic of 15.8 and corrosponding p-value well under 0.05 implies that the model without the interaction term perfroms significantly worse. As such, we conclude that X1^2 should not be dropped.
```{r}
NH <- lm(Y ~ X2 + I(X2^2))
AH <- lm(Y ~ X1 + I(X1^2) + X2 + I(X2^2) + X1:X2)
anova(NH, AH)
```
The significant F-statistic of ~18.1 (p-value << 0.05) implies there is sufficient evidence to reject the null hypothesis.