---
title: "Homework4"
author: "Ryan Peroutka"
date: "March 27, 2018"
output: pdf_document
---
# 5.3
```{r, include=FALSE}
library(alr4)
library(lsmeans)
```
## 1
```{r}
attach(UN11)
m1 <- lm(lifeExpF ~ group)
lsmeans(m1, pairwise ~ group)
```
The results in the second table line up with those presented in the table 5.2 in the book.

## 2
```{r}
m1 <- lm(lifeExpF ~ group + log(ppgdp))
lsmeans(m1, pairwise ~ group)
```
As a result of including log(ppgpd), the difference between oecd and other appears to be significantly different. oecd - africa and other - africa do not appear to have been affected.

#5.4
## 1
```{r}
attach(MinnLand)
boxplot(log(acrePrice)~year)
```
The pattern of prices decreasing starting in 2007 does not appear to be the case with respect to farm sales. Although, they do start to drop off after 2010.> x <- MinnLand$acrePrice # Store acrePrice in x from MinnLand

## 2
```{r}
m1 <- lm(log(acrePrice) ~ factor(year))
summary(m1)
```
The intercept, 7.27 represents the average acrePrice in 2002. The following esitimates (2003-2011) represent the diference between that year's mean and 2002's mean. Apart from 2003, all of the t values appear significant. However, these values are only telling us that the means are significantly different from the reference mean - which isn't all that informative.

## 3
```{r}
m2 <- update(m1,~.-1)
summary(m2)
for (yr in 2002:2011){
  print(mean(log(acrePrice[year == yr])))
}
```
It is clear that the printed values are the same as the estimates from the model.
```{r}

for (yr in 2002:2011){
  sd <- sd(log(acrePrice[year == yr]))
  n <- length(acrePrice[year == yr])
  print(sd/sqrt(n))
}
```
It is clear that the printed values differ from the std error values in the model. This is because the calculation for standard error depends on the difference between the mean/value in relation to the mean.

# 5.8
## 1
```{r}
attach(cakes)
m1 <- lm(Y ~ X1 + X2 + I(X1^2) + I(X2^2)+ X1:X2)
summary(m1)
```
## 2
```{r}
m1 <- lm(Y ~ X1 + X2 + I(X1^2) + I(X2^2)+ X1:X2 + block + block*X1 + block*X2)
summary(m1)
```
block1 doesn't seem to have a significant effect. The same is the case for block1's interaction with X2. However, the interaction between X1 and block1 is significant. 

# 5.10
## 1
(b) accounts for potential interaction between year and region while (a) does not.

## 2
```{r}
m1 <- lm(log(acrePrice) ~ year + region + year:region)
summary(m1)
plot(effect("year:region",m1))
```
In the NorthWest, prices are lowest. As the years progress, the price in the SouthWest rises significantly relative to other regions. Other than this, prices tend to rise a similar amount over time. 