---
title: Homework 2
author: Ryan Peroutka
date: 2/13/18
output: pdf_document
---

# 2.15
## 1
```{r, include=FALSE}
library(alr4)
attach(wblake)
```
### Confidence intervals for fish of ages 2,4,6 y/o respectivley
```{r}
wblakeRegression <- lm(Length~Age)
for (i  in c(2,4,6)) {
  z <- predict(wblakeRegression, data.frame(Age=i), se.fit = TRUE)
  CI.se <- z$se.fit
  alpha <- qt((1-.95)/2, df=z$df)
  print(z$fit + c(alpha, -alpha) * CI.se)
}
```
## 2
### Confidence interval for fish of age 9 y/o
```{r}
z <- predict(wblakeRegression, data.frame(Age=9), se.fit = TRUE)
CI.se <- z$se.fit
print(z$fit + c(alpha, -alpha) * CI.se)
```
There is no data for fish of age 9 and minimal data for fish of age 8.\newline
Therefore, we have no basis to assume the the trend will coninue in the same linear fashion as before.

# 2.20
##  1
```{r, include=FALSE}
attach(oldfaith)
```
```{r}
oldfaithRegression <- lm(Interval~Duration)
intercept <- coef(oldfaithRegression)[1]
slope <- coef(oldfaithRegression)[2]
print(intercept)
print(slope)
```

### Prediction equation
Interval = 0.177 * Duaration + 33.988
\newline\newline
Our the model predicts that, for ever 1 second increase in the duration an eruption, the avergae time ot the next eruption will increase by 0.177 minutes (10.62 seconds).

## 2
### 95% Confidence Interval
```{r}
prediction <- predict(oldfaithRegression, data.frame(Duration=250), se.fit=TRUE) 
sigma.hat <-sqrt(deviance(oldfaithRegression)/df.residual(oldfaithRegression))
SXX <- (sum((Duration - mean(Duration))^2))
PI.se <- sigma.hat * sqrt(1 + 1/df.residual(oldfaithRegression) + (250 - mean(Duration))^2/SXX)
alpha <- qt((1-.95)/2, df=prediction$df)
print(prediction$fit + c(alpha, -alpha) * PI.se)
```
## 3
### 95% Prediction Interval for .9 quantile
```{r}
mean.estimate <- intercept + slope * 250
dev.estimate <- prediction$residual.scal
quantile <- .8 # becuase it's two-sided
x <- quantile * dev.estimate + mean.estimate
CI.se <- prediction$se.fit
alpha <- qt((1-.95)/2, df=prediction$df)
print(x + c(alpha, -alpha) * CI.se)
```
# 3.1

Of the top 6 most outling points, (Afghanistan, East Timor, Mali, Niger, Somalia, Zambia), 4 are in Africa. Taken as a whole, these outliers have significantly lower lifeExpF and pctUrban than the rest of the data.

# 3.3
## 1
```{r, include=FALSE}
attach(BGSgirls)
```
``` {r}
plot(BGSgirls[,c(2,4,1,3,6,11)])
```

The strongest correlations appear to be between WT2/WT9, HT9/ST9 and WT9/BMI18. Some other factors appear to be related, but not quite as strongly.
```{r}
cor(BGSgirls[,c(2,4,1,3,6,11)])
```

The correlation matrix shows failry decent correlations between WT2/WT9, HT9/ST9 and WT9/BMI18. In addition, it shows that HT9/WT9 are strongly correlated, something I failed to notice before.
## 3
```{r}
mlr <- lm(BMI18 ~ HT2 + HT9 + WT2 + WT9 + ST9)
residualStandardError <- 2.14
SYY <- sum((BMI18 - mean(BMI18))^2)
RSS <- (residualStandardError^2) * 64
rSquared <- (SYY - RSS) / SYY
print(residualStandardError)
print(rSquared)
```

```{r}
betaJ <- coef(mlr)
betaJ.se <- coef(summary(mlr))[,2]
betaJ.tStar <- betaJ / betaJ.se
```
### beta1
NH: beta1 = 0
AH: beta1 != 0
```{r}
2*pt(betaJ.tStar[2], df = 64, lower.tail = T)
```
We fail to reject the null hypothesis since we do not have enough evidence to conclude that beta1 is significantly different from 0.

### beta2
NH: beta2 = 0
AH: beta2 != 0
```{r}
2*pt(betaJ.tStar[3], df = 64, lower.tail = F)
```
We fail to reject the null hypothesis since we do not have enough evidence to conclude that beta2 is significantly different from 0.

### beta3
NH: beta3 = 0
AH: beta3 != 0
```{r}
2*pt(betaJ.tStar[4], df = 64, lower.tail = T)
```
We fail to reject the null hypothesis since we do not have enough evidence to conclude that beta3 is significantly different from 0.

### beta4
NH: beta4 = 0
AH: beta4 != 0
```{r}
2*pt(betaJ.tStar[5], df = 64, lower.tail = F)
```
We reject the null hypothesis and conclude that beta4 != 0.

### beta5
NH: beta5 = 0
AH: beta5 != 0
```{r}
2*pt(betaJ.tStar[6], df = 64, lower.tail = T)
```
We reject the null hypothesis and conclude that beta5 != 0.